Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-12-13 ms.tex
Replacing Eq. with Equation

\author[0000-0002-6040-8281]{B. Connor McClellan}
\address{} %TODO: McCormick address
\author[0000-0001-7488-4468]{Shane Davis}
\author[0000-0001-5611-1349]{Phil Arras}
\affiliation{Department of Astronomy, University of Virginia, Charlottesville, VA 22904, USA}
%TODO: Complete postal address for each author, including zip code

@@

Appendix \ref{app:rteqn_derivation} contains a derivation of the transfer equation for convenience. Starting with the full transfer equation, Eq. \ref{eq:finaleqn}, ignoring photon destruction and including a photon emission term given by Eq. \ref{eq:jem}, the steady-state transfer equation is
\be
\nabla^2 J + \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2} & = & 
- \frac{ \sqrt{6} kL}{4\pi \Delta^2} \delta^3(\vec{x} - \vec{x}_s) \delta (\sigma - \sigma_{\rm s}).
\label{eq:rt_no_destr}
\ee
where $J$ is the mean intensity, the spatial variable is $\vec{x}$, and $\vec{x}_s$ is the position of the source. We will consider only the case where $\vec{x}_s=0$. Following \citet{1973MNRAS.162...43H}, we have used a change of variables from $x$ to $\sigma$,
\be \label{eq:int_change_of_variables}
\sigma(x) = \sqrt{\frac{2}{3}}\int_0^x \frac{dx}{\phi(x) \Delta} \approx \sqrt{\frac{2}{3}}\frac{\pi}{a}\frac{x^3}{3}, 
\ee
where the approximation is applicable on the damping wing. The line profile is then approximately (Equation \ref{eq:app:line_profile_wing})
\be \label{eq:line_profile_approx}
\phi \approx \frac{a}{\pi x^2 \Delta} \approx \frac{1}{3 \Delta}\left(\frac{2a}{\pi}\right)^{1/3}|\sigma|^{-2/3}.
\ee
In Equation \ref{eq:rt_no_destr}, $\sigma_s \equiv \sigma(x_s)$ is the photon frequency of the source. $\sigma_s$ is interchangeable with $x_s$ and $\nu_s$ in Doppler widths or Hz, respectively. Balancing the two terms on the left hand side of Equation \ref{eq:rt_no_destr} gives $\sigma \ {\sim}\ \tau_0$, or $x_{\rm peak}\ {\sim}\ (a\tau_0)^{1/3}$. The boundary condition of no incoming intensity at the surface \citep{1986rpa..book.....R} is
\be
J & = & \sqrt{3} H
\label{eq:bc}
\ee
at $r=R$.

@@

A better approximation to the true solution has been derived by \citet{2006ApJ...649...14D}, who generalized the closed-form solution in slab geometry found in \citet{1990ApJ...350..216N}. It satisfies a $J=0$ boundary condition at $r=R$. Again, we generalize their solution to allow emission at frequency $\sigma_{\rm s}$ away from line center. The result can be written as a sum over spatial modes,
\be \label{eq:J0_sum}
J_0 = \frac{\sqrt{6}L}{16\pi \Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\frac{\sin{\kappa_n r}}{\kappa_n r}\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
and
\be \label{eq:H0_sum}
H_0 = - \frac{1}{3k\phi} \frac{\partial J_0}{\partial r} = -\frac{1}{3k\phi}\frac{\sqrt{6}L}{16\pi\Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\left(\frac{\cos{\kappa_n r}}{r} - \frac{\sin{\kappa_n r}}{\kappa_n r^2}\right)\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
where $\kappa_n=n\pi/R$. These can be summed to give the closed form expressions
\be
J_0 & = & \frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right)
\label{eq:J0}
\ee
and
\be
H_0 & = &\frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr^2}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right. \nonumber \\ & & \left. - \left( \frac{\pi r}{R} \right)
\frac{ \cos(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
+ \left( \frac{\pi r}{R} \right)
\frac{ \sin^2(\pi r/R) }{ \left[ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R) \right]^2 }
\right).
\label{eq:H0}
\ee
These solutions agree with Equations \ref{eq:Jd} and \ref{eq:Hd} when the arguments of the trigonometric and hyperbolic functions are small. Again $J_0 \gg H_0$, except near $r=R$, where it goes to zero. The flux at $r=R$ can be written
\be
\nonumber
H_0(R, \sigma) & = & - \frac{1}{3k\phi}
\frac{\sqrt{6}L}{16\pi \Delta}
\frac{1}{R^3}
\sum_{n=1}^{\infty} 
n (-1)^n \exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)}\\
& = &  \frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi \Delta}
\frac{1}{R^3}
\left( 
\frac{ 1 }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] +1 }
\right).
\label{eq:H0surf}
\ee
Eq. \ref{eq:H0surf} will be shown to be a better approximation to the solution than Eq. \ref{eq:Hd}. It is still valid near the delta function at $r=0$, but is also a better approximation at $r=R$. $J_0$ decreases exponentially rather than as a power-law in frequency as it does for $J_d$, giving a much smaller flux on the line wings as compared to the divergent solution.

In order to enforce the boundary conditions, a different solution method is attempted here, namely a continuous Fourier expansion in the frequency variable $\sigma$. The solution of this problem is split into two pieces: $J_0$ which includes the delta function source and satisfies $J=0$ at $r=R$, and $J_{\rm bc}$ which allows the boundary condition $J=\sqrt{3}H$ to be satisfied at $r=R$. The total solution is
\be
J(r,\sigma) & = & J_0(r,\sigma) + J_{\rm bc}(r,\sigma)
\ee
and
\be \label{eq:totalflux}
H(r,\sigma) & = & H_0(r,\sigma) + H_{\rm bc}(r,\sigma).
\ee
The additional term $J_{\rm bc}$ must then be a solution of the homogeneous equation
\be \label{eq:diffeq}
\frac{\partial^2J_{\rm bc}}{\partial r^2} + \frac{2}{r} \frac{\partial J_{\rm bc}}{\partial r}
+ \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J_{\rm bc}}{\partial \sigma^2} &= & 0
\ee
with no delta function source term, and it must allow the boundary conditions to be satisfied at the surface. Since $J_0(R,
\sigma)=0$, the surface boundary condition becomes
\be
J_{\rm bc}(R,\sigma) - \sqrt{3} H_{\rm bc}(R,\sigma) & = 
\sqrt{3} H_0(R,\sigma).
\label{eq:bc2}
\ee
Inserting a frequency dependence $J_{\rm bc} \propto e^{is\sigma}$, for ``wavenumber" $s$, gives the equation for modified spherical Bessel functions of the first kind, $i_0(z)=\sinh(z)/z$ for the radial dependence. The solution can then be represented as
\be
J_{\rm bc}(r,\sigma) & = & 
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\frac{i_0(krs/\Delta)}{i_0(kRs/\Delta)},
\label{eq:Jbc}
\ee
where $A(s)$ is the Fourier amplitude. Inserting Eq. \ref{eq:Jbc} into Eq. \ref{eq:bc2} leads to the following equation for the Fourier amplitudes,
\be
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s)
\left[ 1 + \left( \frac{s}{\sqrt{3} \Delta \phi} \right) \left( \frac{i_0^\prime(kRs/\Delta)}{i_0(kRs/\Delta)} \right) \right]
& = & \sqrt{3} H_0(R,\sigma).
\label{eq:bc3}
\ee
Discretization of Eq. \ref{eq:bc3} for frequency variables $\sigma_i$ and wavenumbers $s_j$
leads to a set of coupled linear equations for the $A(s_j)$. We use equally-spaced points $\delta \sigma = 2\sigma_{\rm max}/(N-1)$ and $\delta s = 2\pi/(N\delta \sigma)$, where $N$ is the number of points for each grid. The result is a linear system to solve for the $A(s)$. The maximum frequency is set as $\sigma_{\rm max} = {\rm constant} \times \tau_0$, for a large enough constant that the end of the frequency grid is at such small intensities that it does not affect the solution except close to the boundaries. The number of points was increased until the solution was well-resolved near line center, and only became inaccurate close to the boundaries. Given the Fourier amplitudes $A(s)$, $J_{\rm bc}$ is computed using Eq. \ref{eq:Jbc}, and the flux is given by
\be
H_{\rm bc}(r,\sigma) & = & -\frac{1}{3k\phi}
\frac{\partial J_{\rm bc}(r,\sigma)}{\partial r}
= -\frac{1}{3k\phi}
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\left( \frac{ks}{\Delta} \right) 
\left( \frac{i_0^\prime(krs/\Delta)}{i_0(kRs/\Delta)} \right).
\label{eq:Hbc}
\ee
The Bessel functions are finite at the center and rise steeply toward the surface when $kRs/\Delta \gg 1$.

@@

In Figure \ref{fig:sol_mc_tau} we show the solutions alongside Monte Carlo for $\tau_0=10^5, 10^6$, and $10^7$. From Eq. \ref{eq:hbc_scaling}, the size of the term $H_{\rm bc}$ should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. Indeed, agreement between Eq. \ref{eq:H0surf} and the Monte Carlo points in Fig. \ref{fig:sol_mc_tau} improves as $\tau_0$ increases, with $H_{\rm bc}$ providing a fractionally smaller correction to $H_0$. One factor of $(a\tau_0)^{1/3}$ has been scaled out of the x-axis such that the peaks of the distributions are horizontally aligned. This scaling has also been applied to the y-axis to preserve normalization of the escape probability. At lower $\tau_0$, the scattering of photons within the Doppler core of the line becomes important, but our analytic solution does not include this effect. The effects of line core scattering can be seen in the Monte Carlo data for $\tau_0=10^5$ and, to a lesser extent, $
\tau_0=10^6$.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{xinit.pdf}
    \caption{$P_{\rm bc}(x) = 16\pi^2R^2H_{\rm bc}(R, x)\Delta/L$ (Eq. \ref{eq:prob_spectrum} with $H \rightarrow H_{\rm bc}$) vs. frequency $x$. Thicker lines indicate larger optical depth labeled by $\tau_0=10^5, 10^7,$ and $10^9$, and darker lines indicate $\sigma_s$ further from line center, labeled by $\sigma_s=0.0, \tau_0, 2\tau_0$. The location of the core-wing boundary $x_{cw}$ is shown with green vertical lines of increasing thickness for each $\tau_0$. The axes are scaled to show that the size of the correction factor agrees with the predicted $(a\tau_0)^{-1/3}$ scaling.}
    \label{fig:xinit}
\end{figure}

Figure \ref{fig:xinit} shows how the correction $P_{\rm bc}(x)$ (Eq. \ref{eq:prob_spectrum} with $H \rightarrow H_{\rm bc}$) scales with both $\tau_0$ and $x_s$. In this figure, $\sigma_s$ is shifted by integer multiples of $\tau_0$ (Eq. \ref{eq:change_of_variables}) such that the source falls near the peak of the spectrum for each $\tau_0$. For clarity, only the $x > 0$ side of the spectrum is shown. From Eq. \ref{eq:hbc_scaling}, it is expected that the fractional size of $H_{\rm bc}$ relative to $H_0$ should become smaller with larger optical depths following $(a\tau_0)^{-1/3}$. This factor has been scaled out of the figure such that solutions for different $\tau_0$ and the same $\sigma_s$ should show close agreement in scale on the figure's vertical axis if the relation holds. Indeed, the scaled solutions converge as $\tau_0$ becomes larger, indicating agreement with the $(a\tau_0)^{-1/3}$ scaling. The remaining discrepancy present in the vertical axis for fixed $\sigma_s$ results from $x_{\rm peak}$ becoming close to $x_{\rm cw}$; at lower $\tau_0$, this causes the line profile approximation in the wing (Eq. \ref{eq:app:line_profile_wing}) to break down. From this, we conclude that the errors introduced by the incorrect separation of variables in \citet{1973MNRAS.162...43H}, \citet{1990ApJ...350..216N}, \citet{2006ApJ...649...14D} and others are indeed proportional to $(a\tau_0)^{-1/3}$.

@@

We seek now to calculate the eigenmodes $J_{nm}(\sigma)$ and eigenfrequencies $\gamma_{nm}$ for a given spatial $n$. These will be labelled by an index $m=1, 2, ...$. To measure the size of the response to detect where resonances occur, we sum the absolute value of $J(n,\sigma,-i\gamma)$ over the array $\sigma$. We call this response $f$, and use the index $j$ to represent the value of the response at discrete points $\gamma_j$ over a range of $\gamma$. In places where $f_j > f_{j-1}$ and $f_j>f_{j+1}$, we have bracketed a resonance that occurs in the interval $(\gamma_{j-1},\gamma_{j+1})$. We then hone in on the eigenvalue before continuing the sweep in $\gamma$. To refine the eigenmode solution, we evaluate $f_{j-1}$, $f_j$, and $f_{j+1}$ at the points $(\gamma_{j-1},\gamma_{j},\gamma_{j+1})$. Assuming the form in Equation \ref{eq:jnsigmaomega_approx}, a guess at the correct eigenvalue $\gamma_{nm}$ can be calculated from
\be
\gamma_{\rm guess} &=& \frac{b\gamma_{j-1} - \gamma_{j+1}}{b - 1},
\ee
where
\be
b &=& \left(\frac{f_{j} - f_{j-1}}{f_{j} - f_{j+1}}\right)\left(\frac{\gamma_{j+1}-\gamma_{j}}{\gamma_{j-1}-\gamma_{j}}\right).
\ee
This is repeated by replacing initial points $(\gamma_{j-1},\gamma_{j},\gamma_{j+1})$ with closer estimates while the size of the response grows as the resonance is approached. After iterating an eigenvalue $\gamma_{nm}$ to convergence, we now find the corresponding eigenfunction $J_{nm}(\sigma)$. We can solve for $J_{nm}(\sigma)$ by comparing two points $\gamma_1$ and $\gamma_2$ near the resonance, cancelling $C(\gamma, \sigma)$ in the difference to find
\be
J_{nm}(\sigma) & \simeq & \frac{ J(n,\sigma,-i\gamma_1)  - J(n,\sigma,-i\gamma_2) }{ 1/(\gamma_{nm}-\gamma_1) - 1/(\gamma_{nm}-\gamma_2)}.
\ee
\begin{figure}
    \centering
    \includegraphics{Jsoln_n1_m100.pdf}
    \caption{Eigenfunctions $J_{nm}(x)$ for the lowest-order spatial eigenmode $n=1$, and $m=1, ..., 100$ with $x_s=0$ and $\tau_0=10^7$. }
    \label{fig:jsoln}
\end{figure}
\noindent The form of a single eigenmode $J_{nm}(\sigma)$ is oscillatory out to some turning point, $\sigma_{\rm tp}$, at which point the function becomes evanescent. The location of the turning point can be found by ignoring the delta-function discontinuity at the source frequency $\sigma_s$ in Eq. \ref{eq:diffusion_plugged_in} and examining the resulting homogeneous differential equation. We obtain
\be \label{eq:wkb_differential_eqn}
\frac{d^2J}{d\sigma^2} & = & \left[ \left( \frac{\kappa_n \Delta }{k} \right)^2 - \frac{3\phi \gamma\Delta^2}{ck}\right] J,
\ee
where the line profile is approximated as in Eq. \ref{eq:app:line_profile_wing}. When the coefficient on the right hand side is positive, exponential growth or decaying evanescent solutions are found. This occurs on the line wings. When the coefficient on the right hand side is negative, oscillatory solutions are found (propagation), which occurs near the line core. The boundary between propagation and evanescence occurs at the turning point, given by
\be \label{eq:sigma_tp}
\sigma_{\rm tp} & = & \sqrt{\frac{2a}{\pi}}\left( \frac{k \gamma}{ \kappa_n^2 c \Delta} \right)^{3/2}
\ee
Thus, to ensure accuracy in each term of Eq. \ref{eq:Jrsigmat}, the bounds of $\sigma$ must be set sufficiently far outside of $\sigma_{\rm tp}$ such that the function is small at the edges. The scale of an $e$-folding in $J_{nm}(\sigma)$ is $k/(\kappa_n \Delta) = \tau_0 / (\sqrt{\pi} n)$, so a grid of $\sigma$ is chosen that spans a large enough number of $e$-foldings that no oscillatory behavior is present at the boundaries of the domain.

>>>>

\author[0000-0002-6040-8281]{B. Connor McClellan}
\author[0000-0001-7488-4468]{Shane Davis}
\author[0000-0001-5611-1349]{Phil Arras}
\address{Department of Astronomy, University of Virginia, Charlottesville, VA 22904, USA}

@@

Appendix \ref{app:rteqn_derivation} contains a derivation of the transfer equation for convenience. Starting with the full transfer equation, Equation \ref{eq:finaleqn}, ignoring photon destruction and including a photon emission term given by Equation \ref{eq:jem}, the steady-state transfer equation is
\be
\nabla^2 J + \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2} & = & 
- \frac{ \sqrt{6} kL}{4\pi \Delta^2} \delta^3(\vec{x} - \vec{x}_s) \delta (\sigma - \sigma_{\rm s}).
\label{eq:rt_no_destr}
\ee
where $J$ is the mean intensity, the spatial variable is $\vec{x}$, and $\vec{x}_s$ is the position of the source. We will consider only the case where $\vec{x}_s=0$. Following \citet{1973MNRAS.162...43H}, we have used a change of variables from $x$ to $\sigma$,
\be \label{eq:int_change_of_variables}
\sigma(x) = \sqrt{\frac{2}{3}}\int_0^x \frac{dx}{\phi(x) \Delta} \approx \sqrt{\frac{2}{3}}\frac{\pi}{a}\frac{x^3}{3}, 
\ee
where the approximation is applicable on the damping wing. The line profile is then approximately (Equation \ref{eq:app:line_profile_wing})
\be \label{eq:line_profile_approx}
\phi \approx \frac{a}{\pi x^2 \Delta} \approx \frac{1}{3 \Delta}\left(\frac{2a}{\pi}\right)^{1/3}|\sigma|^{-2/3}.
\ee
In Equation \ref{eq:rt_no_destr}, $\sigma_s \equiv \sigma(x_s)$ is the photon frequency of the source. $\sigma_s$ is interchangeable with $x_s$ and $\nu_s$ in Doppler widths or Hz, respectively. Balancing the two terms on the left hand side of Equation \ref{eq:rt_no_destr} gives $\sigma \ {\sim}\ \tau_0$, or $x_{\rm peak}\ {\sim}\ (a\tau_0)^{1/3}$. The boundary condition of no incoming intensity at the surface \citep{1986rpa..book.....R} is
\be
J & = & \sqrt{3} H
\label{eq:bc}
\ee
at $r=R$.

@@

A better approximation to the true solution has been derived by \citet{2006ApJ...649...14D}, who generalized the closed-form solution in slab geometry found in \citet{1990ApJ...350..216N}. It satisfies a $J=0$ boundary condition at $r=R$. Again, we generalize their solution to allow emission at frequency $\sigma_{\rm s}$ away from line center. The result can be written as a sum over spatial modes,
\be \label{eq:J0_sum}
J_0 = \frac{\sqrt{6}L}{16\pi \Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\frac{\sin{\kappa_n r}}{\kappa_n r}\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
and
\be \label{eq:H0_sum}
H_0 = - \frac{1}{3k\phi} \frac{\partial J_0}{\partial r} = -\frac{1}{3k\phi}\frac{\sqrt{6}L}{16\pi\Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\left(\frac{\cos{\kappa_n r}}{r} - \frac{\sin{\kappa_n r}}{\kappa_n r^2}\right)\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
where $\kappa_n=n\pi/R$. These can be summed to give the closed form expressions
\be
J_0 & = & \frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right)
\label{eq:J0}
\ee
and
\be
H_0 & = &\frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr^2}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right. \nonumber \\ & & \left. - \left( \frac{\pi r}{R} \right)
\frac{ \cos(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
+ \left( \frac{\pi r}{R} \right)
\frac{ \sin^2(\pi r/R) }{ \left[ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R) \right]^2 }
\right).
\label{eq:H0}
\ee
These solutions agree with Equations \ref{eq:Jd} and \ref{eq:Hd} when the arguments of the trigonometric and hyperbolic functions are small. Again $J_0 \gg H_0$, except near $r=R$, where it goes to zero. The flux at $r=R$ can be written
\be
\nonumber
H_0(R, \sigma) & = & - \frac{1}{3k\phi}
\frac{\sqrt{6}L}{16\pi \Delta}
\frac{1}{R^3}
\sum_{n=1}^{\infty} 
n (-1)^n \exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)}\\
& = &  \frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi \Delta}
\frac{1}{R^3}
\left( 
\frac{ 1 }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] +1 }
\right).
\label{eq:H0surf}
\ee
Equation \ref{eq:H0surf} will be shown to be a better approximation to the solution than Equation \ref{eq:Hd}. It is still valid near the delta function at $r=0$, but is also a better approximation at $r=R$. $J_0$ decreases exponentially rather than as a power-law in frequency as it does for $J_d$, giving a much smaller flux on the line wings as compared to the divergent solution.

In order to enforce the boundary conditions, a different solution method is attempted here, namely a continuous Fourier expansion in the frequency variable $\sigma$. The solution of this problem is split into two pieces: $J_0$ which includes the delta function source and satisfies $J=0$ at $r=R$, and $J_{\rm bc}$ which allows the boundary condition $J=\sqrt{3}H$ to be satisfied at $r=R$. The total solution is
\be
J(r,\sigma) & = & J_0(r,\sigma) + J_{\rm bc}(r,\sigma)
\ee
and
\be \label{eq:totalflux}
H(r,\sigma) & = & H_0(r,\sigma) + H_{\rm bc}(r,\sigma).
\ee
The additional term $J_{\rm bc}$ must then be a solution of the homogeneous equation
\be \label{eq:diffeq}
\frac{\partial^2J_{\rm bc}}{\partial r^2} + \frac{2}{r} \frac{\partial J_{\rm bc}}{\partial r}
+ \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J_{\rm bc}}{\partial \sigma^2} &= & 0
\ee
with no delta function source term, and it must allow the boundary conditions to be satisfied at the surface. Since $J_0(R,
\sigma)=0$, the surface boundary condition becomes
\be
J_{\rm bc}(R,\sigma) - \sqrt{3} H_{\rm bc}(R,\sigma) & = 
\sqrt{3} H_0(R,\sigma).
\label{eq:bc2}
\ee
Inserting a frequency dependence $J_{\rm bc} \propto e^{is\sigma}$, for ``wavenumber" $s$, gives the equation for modified spherical Bessel functions of the first kind, $i_0(z)=\sinh(z)/z$ for the radial dependence. The solution can then be represented as
\be
J_{\rm bc}(r,\sigma) & = & 
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\frac{i_0(krs/\Delta)}{i_0(kRs/\Delta)},
\label{eq:Jbc}
\ee
where $A(s)$ is the Fourier amplitude. Inserting Equation \ref{eq:Jbc} into Equation \ref{eq:bc2} leads to the following equation for the Fourier amplitudes,
\be
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s)
\left[ 1 + \left( \frac{s}{\sqrt{3} \Delta \phi} \right) \left( \frac{i_0^\prime(kRs/\Delta)}{i_0(kRs/\Delta)} \right) \right]
& = & \sqrt{3} H_0(R,\sigma).
\label{eq:bc3}
\ee
Discretization of Equation \ref{eq:bc3} for frequency variables $\sigma_i$ and wavenumbers $s_j$
leads to a set of coupled linear equations for the $A(s_j)$. We use equally-spaced points $\delta \sigma = 2\sigma_{\rm max}/(N-1)$ and $\delta s = 2\pi/(N\delta \sigma)$, where $N$ is the number of points for each grid. The result is a linear system to solve for the $A(s)$. The maximum frequency is set as $\sigma_{\rm max} = {\rm constant} \times \tau_0$, for a large enough constant that the end of the frequency grid is at such small intensities that it does not affect the solution except close to the boundaries. The number of points was increased until the solution was well-resolved near line center, and only became inaccurate close to the boundaries. Given the Fourier amplitudes $A(s)$, $J_{\rm bc}$ is computed using Equation \ref{eq:Jbc}, and the flux is given by
\be
H_{\rm bc}(r,\sigma) & = & -\frac{1}{3k\phi}
\frac{\partial J_{\rm bc}(r,\sigma)}{\partial r}
= -\frac{1}{3k\phi}
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\left( \frac{ks}{\Delta} \right) 
\left( \frac{i_0^\prime(krs/\Delta)}{i_0(kRs/\Delta)} \right).
\label{eq:Hbc}
\ee
The Bessel functions are finite at the center and rise steeply toward the surface when $kRs/\Delta \gg 1$.

@@

In Figure \ref{fig:sol_mc_tau} we show the solutions alongside Monte Carlo for $\tau_0=10^5, 10^6$, and $10^7$. From Equation \ref{eq:hbc_scaling}, the size of the term $H_{\rm bc}$ should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. Indeed, agreement between Equation \ref{eq:H0surf} and the Monte Carlo points in Fig. \ref{fig:sol_mc_tau} improves as $\tau_0$ increases, with $H_{\rm bc}$ providing a fractionally smaller correction to $H_0$. One factor of $(a\tau_0)^{1/3}$ has been scaled out of the x-axis such that the peaks of the distributions are horizontally aligned. This scaling has also been applied to the y-axis to preserve normalization of the escape probability. At lower $\tau_0$, the scattering of photons within the Doppler core of the line becomes important, but our analytic solution does not include this effect. The effects of line core scattering can be seen in the Monte Carlo data for $\tau_0=10^5$ and, to a lesser extent, $
\tau_0=10^6$.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{xinit.pdf}
    \caption{$P_{\rm bc}(x) = 16\pi^2R^2H_{\rm bc}(R, x)\Delta/L$ (Equation \ref{eq:prob_spectrum} with $H \rightarrow H_{\rm bc}$) vs. frequency $x$. Thicker lines indicate larger optical depth labeled by $\tau_0=10^5, 10^7,$ and $10^9$, and darker lines indicate $\sigma_s$ further from line center, labeled by $\sigma_s=0.0, \tau_0, 2\tau_0$. The location of the core-wing boundary $x_{cw}$ is shown with green vertical lines of increasing thickness for each $\tau_0$. The axes are scaled to show that the size of the correction factor agrees with the predicted $(a\tau_0)^{-1/3}$ scaling.}
    \label{fig:xinit}
\end{figure}

Figure \ref{fig:xinit} shows how the correction $P_{\rm bc}(x)$ (Equation \ref{eq:prob_spectrum} with $H \rightarrow H_{\rm bc}$) scales with both $\tau_0$ and $x_s$. In this figure, $\sigma_s$ is shifted by integer multiples of $\tau_0$ (Equation \ref{eq:change_of_variables}) such that the source falls near the peak of the spectrum for each $\tau_0$. For clarity, only the $x > 0$ side of the spectrum is shown. From Equation \ref{eq:hbc_scaling}, it is expected that the fractional size of $H_{\rm bc}$ relative to $H_0$ should become smaller with larger optical depths following $(a\tau_0)^{-1/3}$. This factor has been scaled out of the figure such that solutions for different $\tau_0$ and the same $\sigma_s$ should show close agreement in scale on the figure's vertical axis if the relation holds. Indeed, the scaled solutions converge as $\tau_0$ becomes larger, indicating agreement with the $(a\tau_0)^{-1/3}$ scaling. The remaining discrepancy present in the vertical axis for fixed $\sigma_s$ results from $x_{\rm peak}$ becoming close to $x_{\rm cw}$; at lower $\tau_0$, this causes the line profile approximation in the wing (Equation \ref{eq:app:line_profile_wing}) to break down. From this, we conclude that the errors introduced by the incorrect separation of variables in \citet{1973MNRAS.162...43H}, \citet{1990ApJ...350..216N}, \citet{2006ApJ...649...14D} and others are indeed proportional to $(a\tau_0)^{-1/3}$.

@@

We seek now to calculate the eigenmodes $J_{nm}(\sigma)$ and eigenfrequencies $\gamma_{nm}$ for a given spatial $n$. These will be labelled by an index $m=1, 2, ...$. To measure the size of the response to detect where resonances occur, we sum the absolute value of $J(n,\sigma,-i\gamma)$ over the array $\sigma$. We call this response $f$, and use the index $j$ to represent the value of the response at discrete points $\gamma_j$ over a range of $\gamma$. In places where $f_j > f_{j-1}$ and $f_j>f_{j+1}$, we have bracketed a resonance that occurs in the interval $(\gamma_{j-1},\gamma_{j+1})$. We then hone in on the eigenvalue before continuing the sweep in $\gamma$. To refine the eigenmode solution, we evaluate $f_{j-1}$, $f_j$, and $f_{j+1}$ at the points $(\gamma_{j-1},\gamma_{j},\gamma_{j+1})$. Assuming the form in Equation \ref{eq:jnsigmaomega_approx}, a guess at the correct eigenvalue $\gamma_{nm}$ can be calculated from
\be
\gamma_{\rm guess} &=& \frac{b\gamma_{j-1} - \gamma_{j+1}}{b - 1},
\ee
where
\be
b &=& \left(\frac{f_{j} - f_{j-1}}{f_{j} - f_{j+1}}\right)\left(\frac{\gamma_{j+1}-\gamma_{j}}{\gamma_{j-1}-\gamma_{j}}\right).
\ee
This is repeated by replacing initial points $(\gamma_{j-1},\gamma_{j},\gamma_{j+1})$ with closer estimates while the size of the response grows as the resonance is approached. After iterating an eigenvalue $\gamma_{nm}$ to convergence, we now find the corresponding eigenfunction $J_{nm}(\sigma)$. We can solve for $J_{nm}(\sigma)$ by comparing two points $\gamma_1$ and $\gamma_2$ near the resonance, cancelling $C(\gamma, \sigma)$ in the difference to find
\be
J_{nm}(\sigma) & \simeq & \frac{ J(n,\sigma,-i\gamma_1)  - J(n,\sigma,-i\gamma_2) }{ 1/(\gamma_{nm}-\gamma_1) - 1/(\gamma_{nm}-\gamma_2)}.
\ee
\begin{figure}
    \centering
    \includegraphics{Jsoln_n1_m100.pdf}
    \caption{Eigenfunctions $J_{nm}(x)$ for the lowest-order spatial eigenmode $n=1$, and $m=1, ..., 100$ with $x_s=0$ and $\tau_0=10^7$. }
    \label{fig:jsoln}
\end{figure}
\noindent The form of a single eigenmode $J_{nm}(\sigma)$ is oscillatory out to some turning point, $\sigma_{\rm tp}$, at which point the function becomes evanescent. The location of the turning point can be found by ignoring the delta-function discontinuity at the source frequency $\sigma_s$ in Equation \ref{eq:diffusion_plugged_in} and examining the resulting homogeneous differential equation. We obtain
\be \label{eq:wkb_differential_eqn}
\frac{d^2J}{d\sigma^2} & = & \left[ \left( \frac{\kappa_n \Delta }{k} \right)^2 - \frac{3\phi \gamma\Delta^2}{ck}\right] J,
\ee
where the line profile is approximated as in Equation \ref{eq:app:line_profile_wing}. When the coefficient on the right hand side is positive, exponential growth or decaying evanescent solutions are found. This occurs on the line wings. When the coefficient on the right hand side is negative, oscillatory solutions are found (propagation), which occurs near the line core. The boundary between propagation and evanescence occurs at the turning point, given by
\be \label{eq:sigma_tp}
\sigma_{\rm tp} & = & \sqrt{\frac{2a}{\pi}}\left( \frac{k \gamma}{ \kappa_n^2 c \Delta} \right)^{3/2}
\ee
Thus, to ensure accuracy in each term of Equation \ref{eq:Jrsigmat}, the bounds of $\sigma$ must be set sufficiently far outside of $\sigma_{\rm tp}$ such that the function is small at the edges. The scale of an $e$-folding in $J_{nm}(\sigma)$ is $k/(\kappa_n \Delta) = \tau_0 / (\sqrt{\pi} n)$, so a grid of $\sigma$ is chosen that spans a large enough number of $e$-foldings that no oscillatory behavior is present at the boundaries of the domain.
